<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Visualizing Decision Trees and Forest objects with Revolution R and the RevoEnhancements package}
-->

Visualizing Decision Trees and Forest objects with Revolution R and the RevoEnhancements package
========================================================

Our package `RevoEnhancements` on [github](https://github.com/RevoEnhancements/RevoEnhancements) now contains an enhancement to `RevoScaler` that enables plotting of `rxDtree` and `rxDForest` objects.

First, load the required packages:

```{r setup}
library(RevoScaleR)
library(RevoEnhancements)
library(ggplot2)
rxOptions(
  computeContext=RxLocalParallel(),
  reportProgress=0
)
```


## Plotting decision trees:

The Revolution function to create a decision tree is `rxDTree`.  We use `rxDTree` to fit a partition tree of the `diamonds` data, restricting the maximum tree depth to 3, simply to make the plot stand out better.


```{r rxDTree}
### Create and plot an rxDTree object

frm <- formulaExpand(price ~ ., data=diamonds)
fit <- rxDTree(frm, diamonds, maxDepth=3)
```


Since the package `RevoEnhancements` exports a method for `plot.rxDTree`, we simply have to call `plot(fit)` to visualise the tree: 

```{r rxDTree-plot-1, fig.width=6, fig.height=4}
par(mar=c(0.2, 0.2, 1.5, 0.2))
par(mai=rep(0.1, 4))
plot(fit)
```

The function `plot.rxDTree` takes additional arguments to control the plot and text characteristics:

* `plotArgs` should be a list and gets passed to `plot.rpart`.  Use this to control plot axes, titles, etc.
* `textArg` is also a list and gets passed to `text.rpart`. Use this to control the text on the plot, e.g. colour (`col`) and character size (`cex`).

For example, to set the text colour to blue and the character expansion to 0.7, use the `textArgs` argument:

```{r rxDTree-plot-2, fig.width=6, fig.height=4}
plot(fit, 
     textArgs=list(col="blue", cex=0.7))
```

To also set the plot title, use the `plotArgs` argument:

```{r rxDTree-plot-3, fig.width=6, fig.height=4}
plot(fit, 
     textArgs=list(col="blue", cex=0.7), 
     plotArgs=list(main="Decision tree"))
```



## Plotting Decision Forests:

For comparison we fit a decision forest on the same `diamonds` data.  This creates a `rxDForest` object. This object is a list and the individual trees are contained in the `forest` element.

```{r rxDForest}
### Create and plot an rxDForest object

fit <- rxDForest(frm, diamonds, maxDepth=3)
forest <- fit$forest
```


Now we can plot the individual trees.  Since each individual tree of an `rxDForest` object is itself and `rxDTree` object, the plotting is identical to what we illustrated earlier:

```{r rxDForest-plot-1, fig.width=6, fig.height=4}
par(mar=c(0.2, 0.2, 1.5, 0.2))
par(mai=rep(0.1, 4))
plot(forest[[1]], textArgs=list(col="blue", cex=0.7))
```


Again, you can take control of the plot with the `plotArgs` and `textArgs` arguments:

```{r rxDForest-plot-2, fig.width=6, fig.height=4}
plot(forest[[1]], 
     textArgs=list(col="blue", cex=0.7), 
     plotArgs=list(main="Decision forest: Tree 1"))
plot(forest[[2]], 
     textArgs=list(col="blue", cex=0.7), 
     plotArgs=list(main="Decision forest: Tree 2"))

```

